---
title: "Gov 50 Final Project"
author: "Sam Davidson"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

# Milestone 4: Drilling down on one variable

For this analysis, I want to use one variable that tracks non-coresidential parental engagement, number of visits per week, to approximate overall how engaged different groups of fathers are. I'm assuming that this variable will be correlated to other engagement variables such as number of meals or going to appointments with children. I am going to be looking at two different groups of fathers in particular, fathers that are divorced and fathers that never married. What we can see from this data is that fathers of children ages 0-4 who are divorced see their kids, on average, about 0.22 days fewer per week than fathers who never married, but fathers of children ages 5-18 who are divorced see their kids, on average, about 0.19 days more per week than fathers who never married. The effect of marriage status is different depending on the ages of the non-coresidential children. 


```{r}
library(tidyverse)
men <- read_csv("data/csv_2017_2019_Male.csv")

visit_04 <- men |> filter(NRVISIT04 != "NA" & NRVISIT04 != "Not ascertained") |> mutate(visit_bool_04 = case_when(NRVISIT04 == "Not at all" ~ 0, NRVISIT04 == "Less than once a week" ~ 0.5, NRVISIT04 == "About once a week" ~ 1, NRVISIT04 == "Several times a week" ~ 3, NRVISIT04 == "Every day (at least once a day)" ~ 7))

visit_518 <- men |> filter(NRVISIT518 != "NA" & NRVISIT518 != "Not ascertained" & NRVISIT518 != "Don't know") |> mutate(visit_bool_518 = case_when(NRVISIT518 == "Not at all" ~ 0, NRVISIT518 == "Less than once a week" ~ 0.5, NRVISIT518 == "About once a week" ~ 1, NRVISIT518 == "Several times a week" ~ 3, NRVISIT518 == "Every day (at least once a day)" ~ 7))

ate_marriage_status_04 <- visit_04 |> group_by(marstat) |> summarize(avg_visits_04 = mean(visit_bool_04)) |> filter(marstat == "Divorced or annulled" | marstat == "Never been married")

ate_marriage_status_518 <- visit_518 |> group_by(marstat) |> summarize(avg_visits_518 = mean(visit_bool_518)) |> filter(marstat == "Divorced or annulled" | marstat == "Never been married")

ate_mar_total <- left_join(ate_marriage_status_04, ate_marriage_status_518)

ate_mar_total


```


# Milestone 3: Data Visualization

```{r fig.height = 5, fig.width = 15, echo=FALSE}
library(tidyverse)
men <- read_csv("data/csv_2017_2019_Male.csv")

men |> filter(ncall != "none") |> group_by(marstat) |> summarize(cur_mar = n()) |> ggplot(aes(y=marstat, x=cur_mar)) + geom_col() + labs(title = "Distribution of current marriage status among fathers reporting at least 1 non-coresidential child", x = "Count", y = "Marriage Status")

```
# Data Source

My dataset is the CDC's National Survey of Family Growth (NSFG), conducted between 2017 and 2019. The specific data I'm looking at is an independent sample of men aged 15-49. 

# Research Question

I am interested in looking at non-coresidential father involvement in their children's lives, I want to then see how different factors such as educational attainment, income, and former marital status impacts that involvement. I will have to aggregate several different measures of child involvement present in the survey in order to have one single dependent variable. 

# Hypothesis

I think that non-coresidential fathers who were previously married would be more involved in their children's lives than fathers who were never married, because they would potentially have a closer or better relationship with their ex. I am interested in seeing if this effect is different across different income spectrums, however. 

# Variable Notes

CDC codebook on fathering: https://www.cdc.gov/nchs/data/nsfg/2017-2019_NSFG_Male_SectionG_Codebook-508.pdf

Variables of interest for coresidential focal child: (pages 1-12)
- ROUTG04: In the last four weeks, how often did you
Spend time with [NAME OF CORESIDENTIAL FOCAL CHILD] on an outing away from home to places such as
museums, zoos, movies, sports, playground, park, etc.?
- RMEAL04: (In the last four weeks, how often did you... )
Eat evening meals together with [NAME OF CORESIDENTIAL FOCAL CHILD]?
- And also: RERRAND04, RPLAY04, RREAD04, RAFFECT04, RPRAISE04, RFEED04,RBATH04, RDIAPER04, RBED04, RAPPT04
- And those variables ending in 518, which signifies a coresidential focal child between 5-18 years of age. 

For non-residential focal child:
- NRVISIT04, NRSATVIS04, NROUTG04, etc. (pages 13-24)
- NRVISITG518, etc.

Codebook on Former Wives and First Cohabiting Partner: https://www.cdc.gov/nchs/data/nsfg/2017-2019_NSFG_Male_SectionE_Codebook-508.pdf

Never married variable: nformwife

Other interesting:
- How did the marriage end: MARREND4 (first former wife) --> MARREND7 (fourth former wife)
- #CASEID - this is how you can individually identify each person

Complications:
- There are overlaps between non-coresidential and coresidential fathers. 
- There are not just never married or divorced non-coresidential fathers. There are also fathers who are still married - the challenge is to disaggregate 


